---
title: "In class Practice L20"
author: "Update your name here"
date: "2026-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Q1:  Write the corresponding assumed regression model symbolically.

**Example: `formula = Balance~Income + Rating + Income:Rating`**
          
**Model**:  $E(y)=\beta_0 + \beta_1\times Income +\beta_2\times Rating + \beta_3 \times Income\times Rating$




* `formula = Balance~(Limit + Rating)^2`

**Model**:  $E(y)=\beta_0 +\beta_1\times Limit +\beta_2\times Rating + \beta_3 \times Limit\times Rating$




* `formula = Balance~(Income + Rating)^2 + I(Income^2) + I(Rating^2)`

**Model**: $E(y)=\beta_0 + \beta_1\times Income +\beta_2\times Rating + (\beta_3\times Income\times Rating) + (\beta_4\times Income)^2+(\beta_5\times Rating)^2 $




* `formula = Balance~(Limit + Rating + Income)^2`

**Model**:  $E(y)=\beta_0+\beta_1\times Limit + \beta_2\times Rating+\beta_3\times Income + \beta_4\times (Limit\times Rating\times Income)$ Wrong, look at next one





* `formula = Balance~(Limit + Rating + Income)^2 + I(Limit^2) + I(Rating^2) + I(Income^2)`

**Model**:  $E(y)=\beta_0+\beta_1\times Limit + \beta_2\times Rating+\beta_3\times Income + \beta_4\times (Limit\times Rating) + \beta_5\times (Rating\times Income)+ \beta_6\times (Income\times Limit)+ \beta_7\times Limit^2 + \beta_8\times Rating^2 + \beta_9\times Income^2$



### Q2: Read data `Credit_Group.csv` into R as `credit`. Create `Model0`. Print summary. Check the significance of the interaction term. 

**Read data**
```{r}
credit <- read.csv("Credit_Group.csv")
head(credit)
```

**Create `Model0`, and print summary**
```{r}
Model0 <- lm(Balance ~ (Income+Cards)^2, data=credit)
summary(Model0)
```

**Check the significance of the interaction term.**

p=.712
There is not a significant interaction between Income and Cards



### Q3: Construct an interaction plot. Describe the relationship between estimated Balance and Income and how it depends on Cards. 

**Create three conditions of Cards: low Cards, medium Cards and high Cards**
```{r}
lowcard <- mean(credit$Cards) - sd(credit$Cards)
medcard <- mean(credit$Cards)
highcard <- mean(credit$Cards) + sd(credit$Cards)
```

**Create three corresponding lines/equations**
```{r}
# Model0$coefficients
b0 <-Model0$coefficients[1] ##Plug these in instead of values
b1 <-Model0$coefficients[2]
b2 <-Model0$coefficients[3]
b3 <-Model0$coefficients[4]

# three corresponding equations
lowbal <- 173.5283 + 5.6749*credit$Income + 23.5993*lowcard + .152*credit$Income*lowcard
medbal <- 173.5283 + 5.6749*credit$Income + 23.5993*medcard + .152*credit$Income*medcard
highbal <-173.5283 + 5.6749*credit$Income + 23.5993*highcard + .152*credit$Income*highcard

```


**Create Interaction plot**
```{r}
     ###x vs y
plot(credit$Income, lowbal, type="l", xlab="Income")
lines(credit$Income, medbal, col="red")
lines(credit$Income, highbal, col="blue")
legend(3,1000, legend=c("Low Cards", "Medium Card", "High Cards"), col=c("black", "red", "blue"), lty=1, cex=.8)
```

**Describe the relationship** 

Cant interpret plot since interaction not significant but;

We can see that the lines are parralell. This shows that the relationship between Balance and Income does not depend on cards. There is no significant interaction between Income and Cards.

When the Income increases so does Balance meaning there is a positive relationship, as cards increase balance but it will not effect the relationship between income and balance.




### Q4: Fit **the complete second order model** `Model1`. Print summary.

```{r}
Model1 <- lm(Balance ~ (Income+Limit+Rating)^2 + I(Income^2)+ I(Limit^2)+ I(Rating^2), data=credit)
summary(Model1)
```



### Q5: Using `Model1`, find a point estimate.

```{r}
predict(Model1, newdata=data.frame(Income=50, Rating=355, Limit=4700))

```



### Q6: Repeat Question 4 using centered variables. Call this model `Model2`.

```{r}
credit$Income.cnt <- credit$Income - mean(credit$Income)
credit$Limit.cnt <- credit$Limit - mean(credit$Limit)
credit$Rating.cnt <- credit$Rating - mean(credit$Rating)

Model2 <- lm(Balance ~ (Income.cnt+Limit.cnt+Rating.cnt)^2 + I(Income.cnt^2)+ I(Limit.cnt^2)+ I(Rating.cnt^2), data=credit)
summary(Model2)
```



### Q7: Why do we center variables before building regression models? Does the centering help to improve the model in this case?

We center variables to account for the interaction term, if the variables are uncentered the coefficients may be unreliable and the p value may not be statistically significant. In this case centering the variables did improve the model since the P values for each of the 3 variables is now significant. 


It often reduce the correlation between the individual variable and the higher order terms





### Q8: Using `Model2`, find a point estimate.  Does your answer change from the one in Question 5.

```{r}

predict(Model2, newdata=data.frame(Income.cnt=50-mean(credit$Income), Rating.cnt=355-mean(credit$Rating), Limit.cnt=4700-mean(credit$Limit)))
```

**Does your answer change from the one in Question 5?**

No


### Q9: Add `Group` into consideration as `model3`, and add group effect to `Model2` to see whether `Group` has a significant effect to the Balance. Hint: remember to change `Group` to factor type. 
   
```{r}

# change `Group` type to factor 
credit$Group <- as.factor(credit$Group)

# add `Group` into regression
Model3 <- lm(Balance ~ (Income.cnt+Limit.cnt+Rating.cnt)^2 + I(Income.cnt^2)+ I(Limit.cnt^2)+ I(Rating.cnt^2)+Group, data=credit)
summary(Model3)
```
   
**Does `Group` have a significant effect to the Balance? How? **

Keeping income, Limit, and Rating fixed, compared to group A group B is normally higher by about 8.008

Yes, the p value for GroupCis 1.58e-11 meaning the is is a significant predictor



### Q10: Using `Model3`, find point estimates for three groups separately. Are predictions support the conclusion in question 9)? 

```{r}
predict(Model2, newdata=data.frame(Income.cnt=50-mean(credit$Income), Rating.cnt=355-mean(credit$Rating), Limit.cnt=4700-mean(credit$Limit), Group=c("A","B","C")))

```

**Are predictions support the conclusion in question 9)? Why? **

Yes, the predictors for group A and B are quite close, no significant difference. But there is significant difference between A and C.
